---
title: "Gun Violence 1982-2023"
output:
  ioslides_presentation: default
  beamer_presentation: default
  slidy_presentation: default
date: "`r Sys.Date()`"
author: "Jennifer Vesper"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo=TRUE, comment="", cache=F)
library(magrittr)
library(leaflet)
library(plotly)
library(ggplot2)
```

```{=html}
<style type="text/css">
body p, div, h1, h2, h3, h4, h5 {
color: black;
font-size: 28px;
font-family: Trebuchet MS;
}

slides > slide.title-slide hgroup h1 {
color: #03060C;
font-family: Trebuchet MS;
background: linear-gradient(to left, #000000, #950000,#CC1313  100%);
background-position: 0 100%;
background-size: 100% 2px;
background-repeat: repeat-x;
}

h2 {
color: #8C1D40; 
}
</style>
```
## 1. Mass Shootings- Year and Number of Fatalities

<p style="font-size: 16px">

$\text{Fatalities}=\beta_0+\beta_1\cdot\text{Year}+\varepsilon; \hspace{1cm}\varepsilon\sim\mathcal{N}(0;\sigma^2)$

</p>

```{r}
# Load ggplot2 package
library(ggplot2)

shootings <- 
  read.csv("~/Desktop/US gun violence/shooting_1982-2023_cleaned.csv")


g <- ggplot(data = shootings, aes(x = year, y = fatalities, 
                                  color = gender)) + 
  geom_point() +
  geom_smooth(aes(group=1), method = "lm", se = FALSE) +
 
  labs(x = "Year", y = "Fatalities") +
  ggtitle("Fatalities in Shootings by Year and Gender")

```

\newpage

## 2. Graph showing year of shooting and number of fatalities

This graph shows almost all mass shooters have been men. It also shows there was an increase in the number of fatalities from \~2007-2022

```{r echo=FALSE, fig.height=4,fig.width=8}
  g
```

## 3. Mass Shootings by race, age, and number of fatalities

```{r echo=FALSE}
# Load ggplot2 package
library(ggplot2)

shootings <- 
  read.csv("~/Desktop/US gun violence/shooting_1982-2023_cleaned.csv")

plot_ly(data = shootings, x = ~race, y = ~age_of_shooter, z = ~fatalities, color = ~gender) %>%
  add_markers() %>%
  layout(scene = list(xaxis = list(title = "Race"), yaxis = list(title = "Age of Shooter"), zaxis = list(title = "Fatalities")))

```

## 4. Mass shootings and mental health

There is a lot of debate on whether we have a gun problem or a mental health problem

```{r echo = FALSE}
library(ggplot2)

ggplot(data = shootings, aes(x = prior_signs_mental_health_issues, 
                             fill = gender)) +
  geom_bar(position = "dodge", stat = "count") +
  ggtitle("Number of Shootings by Gender and Prior Signs of Mental Health Issues")
  labs(x = "Mental Health Issues", 
       y = "Count", fill = "Gender")


```

## 5. Mass Shootings - was the weapon obtained legally?

```{r echo=TRUE}
library(plotly)

plot1 <- plot_ly(data = shootings, x = ~year, y = ~location, 
                 color = ~weapons_obtained_legally,
        type = "scatter", mode = "markers") %>%
  layout(title = "Weapons Obtained Legally by Year and Type",
         xaxis = list(title = "Year", dtick = 10), 
         yaxis = list(title = "Location"))


```

## 6. Graph: location, year, and if the weapon was obtained legally.

<p style="font-size: 10px;text-align:center;">

$\text{Percentage Obtained Legally} =\frac{95}{141}\times{100} =\text{67.4%}$ <br> $\text{Percentage Obtained Illegally} =\frac{16}{141}\times{100} =\text{11.3%}$<br>$\text{Percentage TBD/Unknown} =\frac{30}{141}\times{100} =\text{21.3%}$

</p>

</p>

```{r echo=FALSE}
plot1
```

## 7. Frequency and Age Range

```{r}
library(ggplot2)

shootings <- 
  read.csv("~/Desktop/US gun violence/shooting_1982-2023_cleaned.csv")

shootings_clean <- na.omit(shootings)

g3 <- ggplot(shootings_clean, aes(x = age_group, fill = age_group)) +
  geom_bar(stat = "count", color = "black") +
  labs(title = 
         "Frequency of Shootings by Age Group",
       x = "Age Group", y = "Count")



```

## 8. Histogram Showing Frequency of Shootings Based on Age Range

```{r echo=FALSE}
g3
```
